---
title: "Distance Exercises"
output: html_notebook
---

```{r install}
library(devtools)
install_github("genomicsclass/tissuesGeneExpression")
```

```{r load}
library(tissuesGeneExpression)
data(tissuesGeneExpression)
head(e)
head(tissue)
```

The data represents RNA expression levels for eight tissues, each with several **biological replictes**. We call samples that we consider to be from the same population, such as liver tissue from different individuals, biological replicates:

1. How many biological replicates for hippocampus?

```{r}
table(tissue)
```

*Interesting definition of **biological replicates**.  As always replicates are relevant to the domain you are considering. Could be an organ, person, enthnic group or population of a country.*

2. What is the distance between samples 3 and 45?

```{r}
e1 <- e[,3]
e2 <- e[,45]

dq2 <- sqrt(sum(e1-e2)^2)
```

3. What is the distance between gene 210486_at and 200805_at?

```{r}

```


*Interesting to consider checking the distance between two genes. Gene similarity by clustering.*

