---
title: "Distance Exercises"
output: html_notebook
---

* http://genomicsclass.github.io/book/pages/distance.html
* http://genomicsclass.github.io/book/pages/distance_exercises.html

```{r install}
library(devtools)
install_github("genomicsclass/tissuesGeneExpression")
```

```{r load}
library(tissuesGeneExpression)
data(tissuesGeneExpression)
head(e)
head(tissue)
```

The data represents RNA expression levels for eight tissues, each with several **biological replictes**. We call samples that we consider to be from the same population, such as liver tissue from different individuals, biological replicates:

1. How many biological replicates for hippocampus?

```{r}
table(tissue)
```

*Interesting definition of **biological replicates**.  As always replicates are relevant to the domain you are considering. Could be an organ, person, enthnic group or population of a country.*

2. What is the distance between samples 3 and 45?

```{r}
e1 <- e[,3]
e2 <- e[,45]

dq2 <- sqrt(sum(e1-e2)^2) #there are better ways to do this. See answer to question 3. 
```

3. What is the distance between gene 210486_at and 200805_at?

```{r}
g1 <- e["210486_at",]
g2 <- e["200805_at",]
dq3a <- sqrt(crossprod(g1-g2))

g12 <- e[c("210486_at", "200805_at"),]
d12 <- dist(g12)
dq3b <- as.matrix(d12)[1,2] #alternative way to get the same answer. 
```


*Interesting to consider checking the distance between two genes. Gene similarity by clustering.*

4. If I run the command (don’t run it!):

`d = as.matrix( dist(e) )`

How many cells will this matrix have?

You're calculating the euclidean difference between each gene so 

```{r}
dim(e)[1]
```

22215 x 22215

5. Compute the distance between all pair of samples:

```{r}
 d = dist( t(e) )
length(d)
```
17766

6. Why is the answer to exercise 5 not `ncol(e)^2`?

Because we take advantage of symmetry: only lower triangular matrix is stored thus only `ncol(e)*(ncol(e)-1)/2` values.
